---
title: "Base R"
author: "MG"
date: "11 czerwca 2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## R Converting data.frame

Basic wide to long reshape

```{r cars}
gdi <- read.csv(stringsAsFactors = FALSE, url("https://jozefhajnala.gitlab.io/r/post/data/ESA2010_GDI.csv"))
head(gdi[, 1:7])

gdi_long <- reshape(data = gdi,
                    direction = "long",
                    varying = 2:23)
head(gdi_long)
```

**data** - almost obviously, this is the data.frame we want to reshape

**varying** - names or indices of columns which we want to stack on each other into a single column

*R automatically recognizes the Y and the years that get translated into the time column. This is because the column names are in a format that reshape can guess automatically: [string].[integer], in our case "Y.1996", "Y.1997", etc. It has a lot of benefits to keep this naming convention for column names before reshaping. If the names have a different character between the [string] and the [integer] (for example "something_1996", "something_1997"), one can specify this character with the sep argument (e.g. sep = "_")

## Including Plots

You can also embed plots, for example:

```{r pressure, echo=FALSE}
plot(pressure)
```

Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.
