---
title: "Z-score"
author: "Magda G."
date: "9 sierpnia 2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Standardize (Z-score) a dataframe

Standardize / Normalize / Z-score / Scale

The standardize() function allows you to easily scale and center all numeric variables of a dataframe. It is similar to the base function scale(), but presents some advantages: it is tidyverse-friendly, data-type friendly (i.e., does not transform it into a matrix) and can handle dataframes with categorical data.

```{r}
library(psycho)
library(tidyverse)

summary(iris)

z_iris <- iris %>% 
  psycho::standardize() 

summary(z_iris)
```

But beware, standardization **does not change (and “normalize”) the distribution!**


```{r}
z_iris %>% 
  dplyr::select(-Species) %>% 
  gather(Variable, Value) %>% 
  ggplot(aes(x=Value, fill=Variable)) +
      geom_density(alpha=0.5) +
      geom_vline(aes(xintercept=0)) +
      theme_bw() +
      scale_fill_brewer(palette="Spectral")
```


